### 2019-04-30 papers/haskell19/1-intro.tex
WIP section 2

The main contribution of this paper are:
%
\begin{itemize}
\item We identify some of the most important limitations of performing random
  testing with automatically derived generators using state-of-the-art tools
  (Section \ref{sec:sources}).
\item We present an extensible mechanism for representing random values built
  upon different source constructions using the Data Types \`a la Carte approach
  (Section \ref{sec:representation}).
\item We develop a modular generation schema, exploiting the extensible nature
  of our representations, which are extended to encode information relevant to
  the generation process at the type level (Section \ref{sec:generation}).
\item We present a very simple yet powerful domain specific language for
  describing extensible generators solely based on the types we used to
  represent the desired shape of our random data (Section \ref{sec:type-level}).
\item We provide a Template Haskell tool for automatically deriving all the
  required machinery presented throughout this paper, and evaluate its
  generation performance using three real-world case studies (Section
  \ref{sec:casestudies}).
\end{itemize}

>>>>

The main contribution of this paper are:
%
\begin{itemize}
\item We identify some of the most important limitations of performing random
  testing with automatically derived generators using state-of-the-art tools
  (Section \ref{sec:sources}).
\item We present an extensible mechanism for representing random values built
  upon different source constructions using Data Types \`a la Carte (Section
  \ref{sec:representation}).
\item We develop a modular generation schema, exploiting the composable nature
  of our representations, which are extended to encode information relevant to
  the generation process at the type level (Section \ref{sec:generators}).
\item We present a very simple yet powerful domain specific language for
  describing extensible generators solely based on the types used to represent
  the desired shape of our random data (Section \ref{sec:typelevel}).
\item We provide a Template Haskell tool for automatically deriving all the
  required machinery presented throughout this paper, and evaluate its
  generation performance using three real-world case studies (Section
  \ref{sec:casestudies}).
\end{itemize}

### 2019-04-30 papers/haskell19/2-sources.tex
Section 4
Added contributions
WIP section 2

\section{Random Testing using Hints}
\label{sec:sources}

Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
lacinia purus, et dictum nunc justo sit amet elit.

@@

A value of type |Html| represents either plain text strings, via the |Text| data
constructor; single and paired HTML tags using the |Sing| and |Tag| data
constructors, respectively; and concatenation of ...

@@

list :: [Html] -> Html
list xs = Tag "ul" (foldl1 Join (Tag "li" <$> xs))

@@

\begin{code}
simplify :: Html -> Html
simplify (Join (Text t1) (Text t2))
  = Text (t1 ++ t2)
simplify (Join (Join (Text t1) x) y)
  = simplify (Join (Text t1) (simplify (Join x y)))
simplify (Join x y)
  = Join (simplify x) (simplify y)
simplify (Tag t x)
  = Tag t (simplify x)
simplify x = x
\end{code}

>>>>

\section{Random Testing with Source Hints}
\label{sec:sources}

This section introduces the common workflow for generating random values using
automatically derived QuickCheck generators, along with its common limitations.
%
To illustrate this, let us consider the following data type definition encoding
basic HTML pages:

@@

This type allows us to build HTML pages via four possible data constructors:
|Text| is used for plain text values; |Sing| and |Tag| represent singular and
paired HTML tags, respectively; and |Join| to simply concatenate two HTML pages
one after another.
%
Note that the constructors |Tag| and |Join| are recursive, as they have at least
one field of their own type.
%
In this paper, and in the scope of random generation in general, we will refer
to any non-recursive data constructor as \emph{terminal}, and
\emph{non-terminal} in any other case.

Then, using this representation the example page\linebreak
%
\texttt{<html>hello<hr><b>world!</b></html>}
%
can be encoded simply as:
%
\begin{code}
Tag "html" $  Text "hello"
              iJoin Single "hr"
              iJoin Tag "b" $ Text "world!"
\end{code} %$

Then, in order to generate random values of type |Html|, QuickCheck requires us
to provide a random data generator for it.

@@

list :: [Html] -> Html
list []  = Text "empty list"
list xs  = Tag "ul" $ foldl1 Join (Tag "li" <$> xs)

@@

\begin{code}
  gen d = frequency
    [ ...
    , (1, pure br)
    , (5, bold   <$> gen (d-1))
    , (2, list   <$> listOf (gen (d-1)))
    , (3, (<+>)  <$> gen (d-1) <*> gen (d-1)) ]
\end{code}%$

\begin{code}
simplify :: Html -> Html
simplify (Text t1 iJoin Text t2)
  = Text (t1 ++ t2)
simplify ((Text t1 iJoin x) iJoin y)
  = simplify (Text t1 iJoin simplify (x iJoin y))
simplify (x iJoin y)
  = simplify x iJoin simplify y
simplify (Tag t x)
  = Tag t (simplify x)
simplify x = x
\end{code}

\begin{code}
  gen d = frequency
    [ ...
    , (2, do  t1  <- arbitrary
              t2  <- arbitrary
              return (Text t1 iJoin Text t2))
    , (4, do  t2  <- arbitrary
              x   <- go (d-1)
              y   <- go (d-1)
              return ((Text t1 iJoin x) iJoin y))]
\end{code}%$
